# Missing values


```{r}

library(readxl)
library(tidyverse)
library(ggpubr)
library(ggridges)
library(gapminder)
library(forcats)
library(dplyr)

```

In order to detect the missing values within the dataset we selected, we firstly compare the number of rows within each dataset and the complete rows (rows without missing values or invalid values) within it. 


```{r}

covid_raw_data = read.csv(file = 'COVID-19data/dataset 2/data.csv')

covid_weekly_case = covid_raw_data[covid_raw_data$indicator == 'cases',]
covid_weekly_death = covid_raw_data[covid_raw_data$indicator == 'deaths',]

tidy_covid_weekly_func <- function(covid_weekly,mode){
  if (mode == "cum"){
    id_var = "cumulative_count"
    }
  else{id_var = "weekly_count"
    }

  tidy_weekly = pivot_wider(covid_weekly[,c("country", 
                                            "country_code", 
                                            "continent", 
                                            id_var,
                                            "year_week")], 
                            names_from = year_week, 
                            values_from = id_var,
                            values_fill = 0)
  
  tidy_weekly = tidy_weekly[order(tidy_weekly$continent,
                                  tidy_weekly$country),]
  
  return(tidy_weekly)
}

tidy_case_cum = tidy_covid_weekly_func(covid_weekly_case,"cum")
tidy_case_count = tidy_covid_weekly_func(covid_weekly_case,"count")

tidy_death_cum = tidy_covid_weekly_func(covid_weekly_death,"cum")
tidy_death_count = tidy_covid_weekly_func(covid_weekly_death,"count")

# The four datasets above are final outputs
```

```{r}
col_names = colnames(tidy_case_count)

for (i in 4: length(col_names)){
  date_ls = str_split(col_names[i], "-")
  year = as.numeric(date_ls[[1]][1])
  week = as.numeric(date_ls[[1]][2])-1
  new_date = as.Date(paste(year, week, "4", sep="-"), "%Y-%U-%u")
  col_names[i] = as.character(new_date)
}

colnames(tidy_case_count) = col_names
colnames(tidy_death_count) = col_names

colnames(tidy_case_cum) = col_names
colnames(tidy_death_cum) = col_names

```


```{r}
miss = data.frame(dataset='tidy_case_count', source='covid', num_rows=nrow(tidy_case_count), complete_cases=sum(complete.cases(tidy_case_count)))
miss[nrow(miss) + 1,] = c('tidy_death_count', 'covid', nrow(tidy_death_count), sum(complete.cases(tidy_death_count)))
miss[nrow(miss) + 1,] = data.frame(dataset='tidy_case_cum', source='covid', num_rows=nrow(tidy_case_cum), complete_cases=sum(complete.cases(tidy_case_cum)))
miss[nrow(miss) + 1,] = c('tidy_death_cum', 'covid', nrow(tidy_death_cum), sum(complete.cases(tidy_death_cum)))
```



```{r}

# Load in the data with the first three lines skipped
int_trade_1 = read_excel("ft900xlsx/exh1.xlsx",skip = 4)

col_names = colnames(int_trade_1)

col_names_sub = int_trade_1[1,]

# Combine the first two lines to get new column names
for (i in 2:length(col_names)){
  if (col_names[i] %in% c("Balance", "Exports", "Imports")){
    prefix = col_names[i]
  }
  col_names[i] = paste(prefix, int_trade_1[1,][i])
}
colnames(int_trade_1) = col_names

# Removed the useless rows
int_trade_1 = int_trade_1[-c(1,46,47,48,49,50), ]

row_names = int_trade_1$Period

for (i in 1:length(row_names)){
  if (row_names[i] %in% c("2019", "2020", "2021")){
    prefix = row_names[i]
  }
  row_names[i] = paste(prefix,"-",row_names[i])
}

# Remove the rows with na
int_trade_1$Period = row_names
int_trade_1 = int_trade_1 %>%
  filter(complete.cases(.))

# Final outcome is int_trade_1

miss[nrow(miss) + 1,] = c('int_trade_1', 'exh1', nrow(int_trade_1), sum(complete.cases(int_trade_1)))
```


```{r}

imp_trade_cat = read_excel("ft900xlsx/exh8.xlsx",
                           skip = 5)                  # remove the title and subtitles

col_names = colnames(imp_trade_cat)

col_subnames = imp_trade_cat[1,]

for (i in 2:length(col_names)){
  col_name = col_names[i]
  # Remove the suffix
  if (substr(col_name, 1, 4) =="Year"){
    col_name = "Year-to-Date"
  }
  col_names[i] = paste(col_name,
                    col_subnames[i])
}

# Set the new row names
colnames(imp_trade_cat) = col_names

imp_trade_cat = imp_trade_cat %>%
  filter(complete.cases(.))

imp_trade_cat = imp_trade_cat[!(imp_trade_cat$`Item (1)` =="Total, Balance of Payments Basis"|
                                  imp_trade_cat$`Item (1)` =="Net Adjustments"|
                                  imp_trade_cat$`Item (1)` =="Total, Census Basis"),]

imp_trade_tot = imp_trade_cat[(imp_trade_cat$`Item (1)` =="Foods, feeds, and beverages"|
                                 imp_trade_cat$`Item (1)` =="Industrial supplies and materials"|
                                 imp_trade_cat$`Item (1)` =="Capital goods, except automotive"|
                                 imp_trade_cat$`Item (1)` =="Automotive vehicles, parts, and engines"|
                                 imp_trade_cat$`Item (1)` =="Consumer goods"|
                                 imp_trade_cat$`Item (1)` =="Other goods"),]

cat_list = rep(NA, length(imp_trade_cat$`Item (1)`))

for (i in 1:length(imp_trade_cat$`Item (1)`)){
  if (imp_trade_cat$`Item (1)`[i] =="Foods, feeds, and beverages"|
      imp_trade_cat$`Item (1)`[i] =="Industrial supplies and materials"|
      imp_trade_cat$`Item (1)`[i] =="Capital goods, except automotive"|
      imp_trade_cat$`Item (1)`[i] =="Automotive vehicles, parts, and engines"|
      imp_trade_cat$`Item (1)`[i] =="Consumer goods"|
      imp_trade_cat$`Item (1)`[i] =="Other goods"){
        prefix = imp_trade_cat$`Item (1)`[i]
      }
  else{
    cat_list[i] = prefix
  }
}

imp_trade_cat$Category = cat_list

imp_trade_cat = imp_trade_cat %>%
  filter(complete.cases(.))

# The output datasets in this section are imp_trade_cat and imp_trade_tot

miss[nrow(miss) + 1,] = c('imp_trade_cat', 'exh8', nrow(imp_trade_cat), sum(complete.cases(imp_trade_cat)))
miss[nrow(miss) + 1,] = c('imp_trade_tot', 'exh8', nrow(imp_trade_tot), sum(complete.cases(imp_trade_tot)))
```



```{r}

exp_trade_cat = read_excel("ft900xlsx/exh7.xlsx",
                           skip = 5)                  # remove the title and subtitles

col_names = colnames(exp_trade_cat)

col_subnames = exp_trade_cat[1,]

for (i in 2:length(col_names)){
  col_name = col_names[i]
  # Remove the suffix
  if (substr(col_name, 1, 4) =="Year"){
    col_name = "Year-to-Date"
  }
  col_names[i] = paste(col_name,
                    col_subnames[i])
}

# Set the new row names
colnames(exp_trade_cat) = col_names

exp_trade_cat = exp_trade_cat %>%
  filter(complete.cases(.))

exp_trade_cat = exp_trade_cat[!(exp_trade_cat$`Item (1)` =="Total, Balance of Payments Basis"|
                                  exp_trade_cat$`Item (1)` =="Net Adjustments"|
                                  exp_trade_cat$`Item (1)` =="Total, Census Basis"),]

exp_trade_tot = exp_trade_cat[(exp_trade_cat$`Item (1)` =="Foods, feeds, and beverages"|
                                 exp_trade_cat$`Item (1)` =="Industrial supplies and materials"|
                                 exp_trade_cat$`Item (1)` =="Capital goods, except automotive"|
                                 exp_trade_cat$`Item (1)` =="Automotive vehicles, parts, and engines"|
                                 exp_trade_cat$`Item (1)` =="Consumer goods"|
                                 exp_trade_cat$`Item (1)` =="Other goods"),]

cat_list = rep(NA, length(exp_trade_cat$`Item (1)`))

for (i in 1:length(exp_trade_cat$`Item (1)`)){
  if (exp_trade_cat$`Item (1)`[i] =="Foods, feeds, and beverages"|
      exp_trade_cat$`Item (1)`[i] =="Industrial supplies and materials"|
      exp_trade_cat$`Item (1)`[i] =="Capital goods, except automotive"|
      exp_trade_cat$`Item (1)`[i] =="Automotive vehicles, parts, and engines"|
      exp_trade_cat$`Item (1)`[i] =="Consumer goods"|
      exp_trade_cat$`Item (1)`[i] =="Other goods"){
        prefix = exp_trade_cat$`Item (1)`[i]
      }
  else{
    cat_list[i] = prefix
  }
}

exp_trade_cat$Category = cat_list

exp_trade_cat = exp_trade_cat %>%
  filter(complete.cases(.))

# The output datasets in this section are exp_trade_cat and exp_trade_tot

miss[nrow(miss) + 1,] = c('exp_trade_cat', 'exh7', nrow(exp_trade_cat), sum(complete.cases(exp_trade_cat)))
miss[nrow(miss) + 1,] = c('exp_trade_tot', 'exh7', nrow(exp_trade_tot), sum(complete.cases(exp_trade_tot)))

```



```{r}

# Load in the data with the first three lines skipped
us_trade_good = read_excel("ft900xlsx/exh12.xlsx",skip = 4)

col_names = colnames(us_trade_good)

col_names_sub = us_trade_good[1,]

# Combine the first two lines to get new column names
for (i in 2:length(col_names)){
  if (col_names[i] %in% c("Balance", "Exports", "Imports")){
    prefix = col_names[i]
  }
  col_names[i] = paste(prefix, "-",us_trade_good[1,][i])
}
colnames(us_trade_good) = col_names

# Removed the useless rows
us_trade_good = us_trade_good[-c(1,46,47,48,49,50), ]

row_names = us_trade_good$Period

for (i in 1:length(row_names)){
  if (row_names[i] %in% c("2019", "2020", "2021")){
    prefix = row_names[i]
  }
  row_names[i] = paste(prefix,"-",row_names[i])
}

# Remove the rows with na
us_trade_good$Period = row_names
us_trade_good = us_trade_good %>%
  filter(complete.cases(.))

# The output dataset is us_trade_good

miss[nrow(miss) + 1,] = c('us_trade_good', 'exh12', nrow(us_trade_good), sum(complete.cases(us_trade_good)))
```



```{r}

# Load in the data with the first three lines skipped
us_trade_pic = read_excel("ft900xlsx/exh13.xlsx",skip = 5)

col_names = colnames(us_trade_pic)

col_subnames = c("Period",
                 "Total Balance of Payments Basis", 
                 "Net  Adjustments", 
                 "Total Census Basis")

for (i in 1:4){
  col_names[i] = col_subnames[i]
}

colnames(us_trade_pic) = col_names

div = which(us_trade_pic$`Total Balance of Payments Basis` == "Exports"|
              us_trade_pic$`Total Balance of Payments Basis` == "Imports")

us_trade_pic_export = us_trade_pic[(div[1]+1):(div[2]-1), ]
us_trade_pic_import = us_trade_pic[(div[2]+1):length(us_trade_pic$Period), ] 

row_names = us_trade_pic_export$Period

for (i in 1:length(row_names)){
  if (row_names[i] %in% c("2019", "2020", "2021")){
    prefix = row_names[i]
  }
  row_names[i] = paste(prefix,"-",row_names[i])
}

us_trade_pic_export$Period = row_names

# Remove the rows with na
us_trade_pic_export = us_trade_pic_export %>%
  filter(complete.cases(.))

row_names = us_trade_pic_import$Period

for (i in 1:length(row_names)){
  if (row_names[i] %in% c("2019", "2020", "2021")){
    prefix = row_names[i]
  }
  row_names[i] = paste(prefix,"-",row_names[i])
}

us_trade_pic_import$Period = row_names

us_trade_pic_import = us_trade_pic_import %>%
  filter(complete.cases(.))


miss[nrow(miss) + 1,] = c('us_trade_pic_import', 'exh13', nrow(us_trade_pic_import), sum(complete.cases(us_trade_pic_import)))
```


```{r}

trade_countries_2021 = read_excel("ft900xlsx/exh14.xlsx",skip = 5)

trade_countries_2021[trade_countries_2021 == "(R)"] = NA

# Remove the useless rows
trade_countries_2021 = trade_countries_2021[1:(which(trade_countries_2021$`Item (1)` == "Unidentified Countries (2)")-1), ]

col_subnames = trade_countries_2021[1,]
col_subnames = col_subnames[!is.na(col_subnames)]

trade_countries_2021 = trade_countries_2021[ , colSums(is.na(trade_countries_2021)) <= 1]

trade_countries_2021 = trade_countries_2021 %>%
  filter(complete.cases(.))

col_names = colnames(trade_countries_2021)

for (i in 2:length(col_names)){
  if (col_names[i] %in% c("Balance", "Exports", "Imports")){
    prefix = col_names[i]
  }
  col_names[i] = paste(prefix, col_subnames[i-1])
}

colnames(trade_countries_2021) = col_names

trade_countries_2021 = trade_countries_2021[!(
  trade_countries_2021$`Item (1)` =="Total Balance of Payments Basis"|
    trade_countries_2021$`Item (1)` =="Net Adjustments"|
    trade_countries_2021$`Item (1)` =="Total Census Basis"),]

cont_list = rep(NA, length(trade_countries_2021$`Item (1)`))

for (i in 1:length(trade_countries_2021$`Item (1)`)){
  if (trade_countries_2021$`Item (1)`[i] == "North America"|
      trade_countries_2021$`Item (1)`[i] == "Europe"|
      trade_countries_2021$`Item (1)`[i] == "Pacific Rim Countries"|
      trade_countries_2021$`Item (1)`[i] == "South/Central America"|
      trade_countries_2021$`Item (1)`[i] == "Africa"|
      trade_countries_2021$`Item (1)`[i] == "Other Countries"
      ){
        prefix = trade_countries_2021$`Item (1)`[i]
      }
  else{
    cont_list[i] = prefix
  }
}

trade_countries_2021$Continent = cont_list

trade_countries_2021 = trade_countries_2021 %>%
  filter(complete.cases(.))

miss[nrow(miss) + 1,] = c('trade_countries_2021', 'exh14', nrow(trade_countries_2021), sum(complete.cases(trade_countries_2021)))
```

```{r}

trade_countries_2020 = read_excel("ft900xlsx/exh14a.xlsx",skip = 5)

trade_countries_2020[trade_countries_2020 == "(R)"] = NA

# Remove the useless rows
trade_countries_2020 = trade_countries_2020[1:(which(trade_countries_2020$`Item (1)` == "Unidentified Countries (2)")-1), ]

col_subnames = trade_countries_2020[1,]
col_subnames = col_subnames[!is.na(col_subnames)]

trade_countries_2020 = trade_countries_2020[ , colSums(is.na(trade_countries_2020)) <= 1]

trade_countries_2020 = trade_countries_2020 %>%
  filter(complete.cases(.))

col_names = colnames(trade_countries_2020)

for (i in 2:length(col_names)){
  if (col_names[i] %in% c("Balance", "Exports", "Imports")){
    prefix = col_names[i]
  }
  col_names[i] = paste(prefix, col_subnames[i-1])
}

colnames(trade_countries_2020) = col_names

trade_countries_2020 = trade_countries_2020[!(
  trade_countries_2020$`Item (1)` =="Total Balance of Payments Basis"|
    trade_countries_2020$`Item (1)` =="Net Adjustments"|
    trade_countries_2020$`Item (1)` =="Total Census Basis"),]

cont_list = rep(NA, length(trade_countries_2020$`Item (1)`))

for (i in 1:length(trade_countries_2020$`Item (1)`)){
  if (trade_countries_2020$`Item (1)`[i] == "North America"|
      trade_countries_2020$`Item (1)`[i] == "Europe"|
      trade_countries_2020$`Item (1)`[i] == "Pacific Rim Countries"|
      trade_countries_2020$`Item (1)`[i] == "South/Central America"|
      trade_countries_2020$`Item (1)`[i] == "Africa"|
      trade_countries_2020$`Item (1)`[i] == "Other Countries"
      ){
        prefix = trade_countries_2020$`Item (1)`[i]
      }
  else{
    cont_list[i] = prefix
  }
}

trade_countries_2020$Continent = cont_list

trade_countries_2020 = trade_countries_2020 %>%
  filter(complete.cases(.))

miss[nrow(miss) + 1,] = c('trade_countries_2020', 'exh14a', nrow(trade_countries_2020), sum(complete.cases(trade_countries_2020)))
```


```{r}

countries_quarters = read_excel("ft900xlsx/exh20.xlsx",skip = 4)

div = which(countries_quarters$`Country and Area` == "Balance"|
              countries_quarters$`Country and Area` == "Exports"|
              countries_quarters$`Country and Area` == "Imports")

end_ind = which(countries_quarters$`Country and Area` == "South/Central America")

balance_ind = 2:(div[2]-1)
export_ind = (div[2]+1):(div[3]-1)
import_ind = (div[3]+1):end_ind[3]


countries_quarters_bal = countries_quarters[balance_ind, ]
countries_quarters_exp = countries_quarters[export_ind, ]
countries_quarters_imp = countries_quarters[import_ind, ]

miss[nrow(miss) + 1,] = c('countries_quarters_bal', 'exh20', nrow(countries_quarters_bal), sum(complete.cases(countries_quarters_bal)))
miss[nrow(miss) + 1,] = c('countries_quarters_exp', 'exh20', nrow(countries_quarters_exp), sum(complete.cases(countries_quarters_exp)))
miss[nrow(miss) + 1,] = c('countries_quarters_imp', 'exh20', nrow(countries_quarters_imp), sum(complete.cases(countries_quarters_imp)))
```

```{r}

countries_quarters_goods = read_excel("ft900xlsx/exh20a.xlsx",skip = 4)

div = which(countries_quarters_goods$`Country and Area` == "Balance"|
              countries_quarters_goods$`Country and Area` == "Exports"|
              countries_quarters_goods$`Country and Area` == "Imports")

end_ind = which(countries_quarters_goods$`Country and Area` == "South/Central America")

balance_ind = 2:(div[2]-1)
export_ind = (div[2]+1):(div[3]-1)
import_ind = (div[3]+1):end_ind[3]


countries_quarters_goods_bal = countries_quarters_goods[balance_ind, ]
countries_quarters_goods_exp = countries_quarters_goods[export_ind, ]
countries_quarters_goods_imp = countries_quarters_goods[import_ind, ]


miss[nrow(miss) + 1,] = c('countries_quarters_goods_bal', 'exh20a', nrow(countries_quarters_goods_bal), sum(complete.cases(countries_quarters_goods_bal)))
miss[nrow(miss) + 1,] = c('countries_quarters_goods_exp', 'exh20a', nrow(countries_quarters_goods_exp), sum(complete.cases(countries_quarters_goods_exp)))
miss[nrow(miss) + 1,] = c('countries_quarters_goods_imp', 'exh20a', nrow(countries_quarters_goods_imp), sum(complete.cases(countries_quarters_goods_imp)))
```


```{r}

countries_quarters_svs = read_excel("ft900xlsx/exh20b.xlsx",skip = 4)

div = which(countries_quarters_svs$`Country and Area` == "Balance"|
              countries_quarters_svs$`Country and Area` == "Exports"|
              countries_quarters_svs$`Country and Area` == "Imports")

end_ind = which(countries_quarters_svs$`Country and Area` == "South/Central America")

balance_ind = 2:(div[2]-1)
export_ind = (div[2]+1):(div[3]-1)
import_ind = (div[3]+1):end_ind[3]


countries_quarters_svs_bal = countries_quarters_svs[balance_ind, ]
countries_quarters_svs_exp = countries_quarters_svs[export_ind, ]
countries_quarters_svs_imp = countries_quarters_svs[import_ind, ]

miss[nrow(miss) + 1,] = c('countries_quarters_svs_bal', 'exh20b', nrow(countries_quarters_svs_bal), sum(complete.cases(countries_quarters_svs_bal)))
miss[nrow(miss) + 1,] = c('countries_quarters_svs_exp', 'exh20b', nrow(countries_quarters_svs_exp), sum(complete.cases(countries_quarters_svs_exp)))
miss[nrow(miss) + 1,] = c('countries_quarters_svs_imp', 'exh20b', nrow(countries_quarters_svs_imp), sum(complete.cases(countries_quarters_svs_imp)))
```




```{r, fig.height = 6, fig.width = 8}
miss_new = miss %>% pivot_longer(cols=c(3,4),names_to = "count type") %>% mutate(value=as.numeric(value)) 


library(viridis)

ggplot(miss_new[1:8,], aes(fill=`count type`, y=value, x=dataset)) + 
    geom_bar(position="dodge", stat="identity") +
    ggtitle("Missing Pattern for Covid Data") +
    xlab("") +
    coord_flip() +
    guides(fill = guide_legend(reverse = TRUE)) +
    ylim(c(0,300)) +
    geom_text(aes(label=value),position=position_dodge(0.9),hjust=-0.2)
    

ggplot(miss_new[9:26,], aes(fill=`count type`, y=value, x=dataset)) + 
    geom_bar(position="dodge", stat="identity") +
    ggtitle("Missing Pattern for BOP Data (Trade)") +
    xlab("") +
    coord_flip() +
    guides(fill = guide_legend(reverse = TRUE)) +
    ylim(c(0,200)) +
    geom_text(aes(label=value),position=position_dodge(0.9),hjust=-0.2)
    

ggplot(miss_new[27:44,], aes(fill=`count type`, y=value, x=dataset)) + 
    geom_bar(position="dodge", stat="identity") +
    ggtitle("Missing Pattern for BOP Data (Countries and Areas)") +
    xlab("") +
    coord_flip() +
    guides(fill = guide_legend(reverse = TRUE)) +
    ylim(c(0,40)) +
    geom_text(aes(label=value),position=position_dodge(0.9),hjust=-0.2)
```

According to the graphs we created above, we find that for every dataset we select, there total number of rows matches the number of complete rows. Hence, we could easily concluded that in all data we selected in this assignment are complete without any missing value, as they are all collected from government-operating websites that guarantee comprehensiveness as well as validity. In addition, we do not need to further fulfill the missing values within them via predicted values.
